CiteQ: Analysis on Citation Sentiment: University of Waterloo Computer Science Faculty Case Study

### 2023-12-30 paper/main.tex
paper content done

\usepackage{bm}
\usepackage{algorithm}
\usepackage[noEnd=True]{algpseudocodex}
\usepackage{tikz}
\usepackage{listings}
\usepackage{float}
\floatstyle{plain}
\newfloat{Figure}{t}{myc}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{CiteQ: Analysis on Citation Purposes: University of Waterloo Computer Science Faculty Case Study}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  TODO: Write abstract
\end{abstract}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Citation Purpose, Citation Analysis, Citation Trends, LLM classification}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

The purpose of this paper is to study the citation purposes and its affects
on other scientometric indicators. We aim to answer the following research 
questions:

\begin{enumerate}
    \item[RQ1] Do the citation purposes affect the total number of
    citations a paper or a researcher receives? (i.e. is there a karma effect?)
    \item[RQ2] How does the distribution of citation purposes change over time?
    \item[RQ3] Is there any difference in the distribution of citation purposes
    between same institution researchers and different institution researchers? 
    \item[RQ4] How does the distribution of citation purposes change over the
    duration of a researcher's career?
\end{enumerate}

To answer these questions, we collected the publication and citation data of
the University of Waterloo Computer Science faculty members. We then used
the large language model (LLM) llama2 \cite{touvron2023llama} to classify the citation 
purposes of each paper. We then analyzed this data to answer the research questions.

\section{Related Work}
Our work is related to a large body of research on citation analysis. Abu-Jbara et. al. \cite{abu2013purpose} provides a good starting point for citation purpose analysis. They propose 6 classes of citation purposes: criticizing, comparison, use, substantiating, basis, and neutral. We use these classes as the basis for our analysis.

\section{Methodology}
\subsection{Data Collection}
We used the Semantic Scholar (SS) API to collect the publication and citation data of all current (as of this writing) members of the University of Waterloo Computer Science faculty members. SS provides an API endpoint to retrieve the context of a citation. We used this endpoint to retrieve the citation context of each citation
made by a faculty member or to a faculty member. In total we collected data about 96 faculty members, 9,438 papers, and 697,609 citations. We formed an SQLite database, with the scheme shown in Figure \ref{fig:citeq-scheme}, to facilitate flexible querying of the data.

@@

\subsection{Citation Purpose Classification}
We experimented with a number of LLMs including chatGPT-3.5 \cite{brown2020language}, chatGPT-4.5 \cite{openai2023gpt4}, mistral \cite{jiang2023mistral}, and llama2 \cite{touvron2023llama}. Although chatGPT models performed slightly better in terms of accuracy, we decided to use llama2 with 7 billion parameters due to its low cost. We used 5 RTX4090 GPUs for 10 hours to classify the citation purposes of all 697,609 citations.

We used the following citation purpose classes: criticizing, comparison, use, substantiating, basis, and neutral. The Figure \ref{figure:prompt-template} shows the prompt template we used.

\subsubsection{Evaluation}
We hand labeled 100 citations to evaluate the accuracy of the classification.
The results are shown in Table \ref{table:evaluation}. We achieved an accuracy of 0.82. The accuracy of the classification is not as high as we would like. However, we believe that it is sufficient for our purposes.

\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|l|l|l|}
        \hline
        \textbf{Class/Model} & \textbf{llama2} & \textbf{mistral} & \textbf{GPT-3.5} & \textbf{GPT-4.5} \\ \hline
        Criticizing & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        Comparison & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        Use & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        Substantiating & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        Basis & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        Neutral & 0.75 & 0.75 & 0.75 & 0.75 \\ \hline
        \textbf{Average} & \textbf{0.82} & \textbf{0.82} & \textbf{0.82} & \textbf{0.82} \\ \hline
    \end{tabular}
    \caption{Precision of different LLMs in citation purpose classification}
    \label{table:evaluation}
\end{table}

\begin{figure}
\begin{lstlisting}
"""
The following is a set of citation purpose categories, each category name is followed by a description of the category and an example of a sentence that belongs to this category.

Name: Criticizing
Description: Criticism can be positive or negative. A citing sentence is classified as "criticizing" when it mentions the weakness/strengths of the cited approach, negatively/positively criticizes the cited approach, negatively/positively evaluates the cited source.
Example: Chiang (2005) introduced a constituent feature to reward phrases that match a syntactic tree but did not yield significant improvement.

Name: Comparison
Description: A citing sentence is classified as "comparison" when it compares or contrasts the work in the cited paper to the author's work. It overlaps with the first category when the citing sentence says one approach is not as good as the other approach. In this case we use the first category.
Example: Our approach permits an alternative to minimum error-rate training (MERT; Och, 2003);

Name: Use
Description: A citing sentence is classified as "use" when the citing paper uses the method, idea or tool of the cited paper.
Example: We perform the MERT training (Och, 2003) to tune the optimal feature weights on the development set.

Name: Substantiating
Description: A citing sentence is classified as "substantiating" when the results, claims of the citing work substantiate, verify the cited paper and support each other.
Example: It was found to produce automated scores, which strongly correlate with human judgements about translation fluency (Papineni et al. , 2002).

Name: Basis
Description: A citing sentence is classified as "basis" when the author uses the cited work as starting point or motivation and extends on the cited work.
Example: Our model is derived from the hidden-markov model for word alignment (Vogel et al., 1996; Och and Ney, 2000).

Name: Neutral (Other)
Description: A citing sentence is classified as "neutral" when it is a neutral description of the cited work or if it doesn't come under any of the above categories.
Example: The solutions of these problems depend heavily on the quality of the word alignment (Och and Ney, 2000).

Classify the following in text citation into one of these categories. Only use a single category for each citation. Respond with only a single word, the name of the category.

{Citation Context}
"""
\end{lstlisting}
\caption{Prompt template used for citation purpose classification}
\label{figure:prompt-template}
\end{figure}

\section{Results}
Initial results

\begin{Figure}
  \begin{tikzpicture}
  \begin{axis}[
      symbolic x 
      coords={criticizing, comparison, use, substantiating, basis, neutral},
      scaled ticks=false,
      style={/pgf/number format/1000 sep=},
      xtick=data,
      xticklabel style = {rotate=30,anchor=east},
      ytick distance=50000,
      ylabel=Number of Vulnerabilities,
      point meta={(y/697609)*100},
      nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%}]
      \addplot[ybar,fill=gray] coordinates {
          (criticizing, 560986)
          (comparison, 12122)
          (use,43646)
          (substantiating, 14860)
          (basis, 8057)
          (neutral, 57938)
      };
  \end{axis}
  \end{tikzpicture}
  \caption{Distribution of Vulnerabilities }
  \label{graph:1}
  \end{Figure}

>>>>

\usepackage{bm}
\usepackage{algorithm}
\usepackage[noEnd=True]{algpseudocodex}
\usepackage{tikz}
\usepackage{listings}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{multirow,siunitx}
\floatstyle{plain}
\newfloat{Figure}{t}{myc}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{CiteQ: Analysis on Citation Sentiment: University of Waterloo Computer Science Faculty Case Study}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
While citation count and H-index is the most important metrics for measuring impact in Science, the intention of the given citation is usually overlooked. To examine the affects of citation sentiments on academic success, this paper conducts a large scale empirical study on University of Waterloo Computer Science faculty members' publication data. We show a negative correlation between H-index and the percentage of negative citations given by that researcher. We also show a positive correlation between number of papers published by a researcher and percentage of neutral citations given by that researcher. In addition we discuss the changes in the distribution of citation sentiments over time and the differences between same institution and different institution citation sentiments. We also provide a dataset of 9,438 papers' metadata and 697,609 citations with their citation sentiments for future research.
\end{abstract}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Citation Sentiment, Citation Analysis, Citation Trends, LLM classification}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

The purpose of this paper is to study the citation sentiments and its affects
on citation counts. We aim to answer the following research 
questions:

\begin{enumerate}
    \item[RQ1] Do the citation sentiments affect the total number of
    citations a paper or a researcher receives? (i.e. is there a karma effect?)
    \item[RQ2] How does the distribution of citation sentiments change over the
    duration of a researcher's career?
    \item[RQ3] How did the distribution of citation sentiments change over time?
    \item[RQ4] Is there any difference in the distribution of citation sentiments
    between same institution researchers and different institution researchers? 
\end{enumerate}

We collected the publication and citation data of
the University of Waterloo Computer Science faculty members. We then used
the large language model (LLM) Solar 10.7B \cite{upstage2023solar} to classify the citation 
sentiments of each paper. We then analyzed this data to answer the research questions.

\section{Related Work}
There is a large body of research on citation analysis. Most of these studies use the ACL Anthology \cite{bird2008acl} as their dataset. ACL Anthology is a dataset of NLP papers and their citations. Using the ACL Anthology creates two main differences from our study. First, it provides more contextual metadata about the citations than our dataset, such as the position of the citation in the paper, the section of the paper that the citation is in, etc. Second, it provides a large enough training dataset to train or finetune any kind of model, which our dataset also lacks. \cite{athar2011sentiment, athar2012context, abu2013purpose, hernandez2015citation} are some of the studies that use the ACL Anthology dataset. They are generally focusing on the task of citation sentiment classification itself and they do not analyze the output of this classification. None of the studies we have found use LLMs for classification. Most use sentence or word embeddings.

\cite{yan2020relationship}, \cite{yan2020authors} are some of the studies that analyze the citation sentiment classification output. \cite{yan2020relationship} analyze the relationship between journal impact and the citation sentiment. They use PubMed Central dataset \cite{pubmed} for this study. \cite{yan2020authors} analyze the relationship between Nobel prize winners and their citation sentiments pre and post Nobel prize. They manually gather a 100 paper dataset for this study.

@@

\sisetup{table-format=2.2} 
\begin{table*}
\centering
    \begin{tabular}{|l|S|S|S|S|S|S|S|S|S|S|}           \hline 
\multirow{2}{*}{type}    & \multicolumn{2}{c|}{Positive}    &\multicolumn{2}{c|}{Negative}  &\multicolumn{2}{c|}{Neutral}  &\multicolumn{2}{c|}{Bad Context} &\multicolumn{2}{c|}{\textbf{Average}}  \\   \cline{2-11}
  &{Prec.} & {Rec.} & {Prec.}  & {Rec.} &{Prec.}  & {Rec.} &{Prec.} &{Rec.}  &{Prec.}  & {Rec.}         \\   \hline 
Solar 10.7B& 0.68 &0.41 & 0.67 & \textbf{0.67} & \textbf{0.47} & 0.87 & 0.90 & 0.44 &0.68 & \textbf{0.60} \\ \hline  
GPT-4.5 & \textbf{0.87} & 0.19 & 0.71 & 0.62 & 0.43 & \textbf{0.96} & 0.92 & \textbf{0.52} & \textbf{0.73} & 0.57 \\ \hline
GPT-3.5 & 0.50 & \textbf{0.95} & \textbf{0.80} & 0.50 & 0.36 & 0.21 & \textbf{1.00} & 0.26 & 0.66 & 0.47\\   \hline
mistral 7x8B & 0.70 & 0.38 & 0.71 & 0.56 & 0.41 & 0.84 & 0.60 & 0.26 & 0.60 & 0.51\\ \hline
llama2 7B & 0.39 & 0.75 & 0.60 & 0.375 & 0.34 & 0.25 & \textbf{1.00} & 0.04 & 0.58 & 0.35\\   \hline 
mistral & 0.48 & 0.27 & 0.36 & 0.50 & 0.43 & 0.87 & 0.75 & 0.13 & 0.50 & 0.44\\ \hline
 \end{tabular}
\caption{Evaluation results of the LLMs}
\label{table:evaluation}
\end{table*}

\section{Methodology}
\subsection{Data Collection}
We used the Semantic Scholar (SS) API to collect the publication and citation data of all current (as of this writing) members of the University of Waterloo Computer Science faculty members. SS provides an API endpoint to retrieve the context of a citation. We used this endpoint to retrieve the citation context of each citation
made by a faculty member or to a faculty member. In total we collected data about 96 faculty members, 9,438 papers, and 697,609 citations. We formed an SQLite database, with the scheme shown in Figure \ref{fig:citeq-scheme}, to facilitate flexible querying of the data. This database is publicly available at \url{https://github.com/onurerenarpaci/citeQ.git}.

\subsection{Citation Sentiment Classification}
We experimented with a number of LLMs including chatGPT-3.5 \cite{brown2020language}, chatGPT-4.5 \cite{openai2023gpt4}, mistral and mistral 7x8B \cite{jiang2023mistral}, llama2 \cite{touvron2023llama} and Solar 10.7B \cite{upstage2023solar}. Among these, only promising results were obtained from Solar 10.7B and \cite{openai2023gpt4}. Since with the size of the dataset we had, we could not afford to run the classification on GPT-4.5, so we chose Solar 10.7B as our LLM.

We used the following citation purpose classes: positive, neutral, negative, bad context. The positive class includes citations that are talking about the strengths of the paper, important it is. The negative class includes citations that are talking about the weaknesses of the paper, or talks about their solution being better than the paper's solution. The neutral class includes citations that are talking about the paper without any positive or negative sentiment. The bad context class is for badly captured citation contexts that do not contain any information about the citation sentiment. With the satisfactory precision on the bad context class, we removed all the citations labeled as bad context from our dataset for all the analysis we have conducted. You can see the prompt template we used in Appendix \ref{appendix:classification-prompt}.

We used 6 Nvidia A100 80GB GPUs to run the classification. The total inference time was 1 day and 4 hours.

\subsubsection{Evaluation}
We hand labeled 100 citations to evaluate the accuracy of the classification.
The results are shown in Table \ref{table:evaluation}. We can see that the Solar 10.7B has the highest average recall, and second highest average precision. We can also see that the Solar 10.7B has the most consistent scores among classes, which is important for our analysis. Therefore, we chose Solar 10.7B as our LLM.

\section{Results}

\begin{Figure}
  \begin{tikzpicture}
  \begin{axis}[
      symbolic x 
      coords={Positive, Neutral, Negative, Bad Context},
      scaled ticks=false,
      style={/pgf/number format/1000 sep=},
      xtick=data,
      xticklabel style = {rotate=30,anchor=east},
      ytick distance=50000,
      ylabel=Number of Vulnerabilities,
      point meta={(y/697609)*100},
      nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%}]
      \addplot[ybar,fill=gray] coordinates {
          (Positive, 150639)
          (Neutral, 457463)
          (Negative, 59054)
          (Bad Context, 30453)
      };
  \end{axis}
  \end{tikzpicture}
  \caption{Distribution of citation purposes}
  \label{graph:distribution}
\end{Figure}

\begin{figure*}
  \centering
  \begin{subfigure}{0.3\textwidth}
    \begin{tikzpicture}
      \begin{axis}[
          width=(1\textwidth),
          xlabel={Negative Citation percentage [\%]},
          ylabel={H-index},
          xmin=0, xmax=30,
          ymin=0, ymax=80,
          xtick={0, 5, 10, 15, 20, 25, 30},
          ytick={0,10,20,30,40,50,60,70,80},
          ymajorgrids=true,
          grid style=dashed,
        ]

        \addplot[
          only marks, color=blue, mark size=1.3pt
          ]
          coordinates {
            (15.15,16)(10.55,16)(14.44,5)(7.21,23)(9.57,23)(7.83,22)(18.18,7)(8.52,18)(6.66,25)(3.92,11)(9.46,32)(6.8,26)(4.01,18)(11.46,66)(7.98,40)(7.73,5)(11.38,23)(12.7,14)(16.05,13)(6.53,24)(7.17,46)(5.53,39)(8.88,31)(10.06,39)(11.83,17)(5.42,16)(8.27,20)(5.59,12)(4.55,23)(10.45,40)(15.78,50)(14.08,19)(12.64,17)(7.23,22)(10.09,10)(6.36,9)(14.99,15)(14.88,33)(7.43,43)(6.77,7)(7.47,45)(8.47,26)(5.62,23)(4.71,31)(14.61,16)(14.2,36)(5.3,31)(9.81,23)(4.1,22)(7.8,16)(7.02,30)(8.63,39)(9.7,26)(11.65,14)(6.63,77)(13.58,8)(6.42,8)(3.52,22)(10.21,7)(7.27,13)(7.69,25)(17.37,14)(13.87,34)(12.28,34)(10.97,8)(5.24,23)(5.61,15)(6.22,20)(10.06,52)(6.38,45)(6.13,20)(11.54,18)(6.08,31)(6.52,18)(4.33,31)(4.03,38)(3.03,10)(7.61,5)(3.53,25)(10.42,22)(15.11,10)(8.56,30)(8.61,16)(8.67,33)(12.79,30)(7.81,13)(4.99,25)(6.04,20)(10.93,20)(11.32,2)(8.76,29)(23.61,8)(9.54,26)(19.64,6)(6.27,19)(5.71,19)
          };
          \addlegendentry{Researchers}

        \addplot[
          mark=square,
          ]
          coordinates {
            (3.33,26.55)(6.67,24.72)(10.0,22.88)(13.33,21.05)(16.67,19.22)(20.0,17.38)(23.33,15.55)(26.67,13.71)
          };
          \addlegendentry{Least Squares Fit}

      \end{axis}
    \end{tikzpicture}
    \caption{Negative citation percentage vs. H-index}
    \label{graph:negative-karma}
\end{subfigure}
\hfill
\begin{subfigure}{0.3\textwidth}
  \begin{tikzpicture}
    \begin{axis}[
        width=(1\textwidth),
        xlabel={Positive Citation percentage [\%]},
        ylabel={H-index},
        xmin=13, xmax=36,
        ymin=0, ymax=80,
        xtick={10, 15, 20, 25, 30, 35},
        ytick={0,10,20,30,40,50,60,70,80},
        ymajorgrids=true,
        grid style=dashed,
    ]

    \addplot[
      only marks, color=blue, mark size=1.3pt
      ]
      coordinates {
        (14.41,16)(17.0,16)(17.78,5)(24.68,23)(19.13,23)(19.33,22)(22.76,7)(28.28,18)(19.87,25)(29.9,11)(20.85,32)(19.76,26)(26.18,18)(20.34,66)(24.02,40)(22.5,5)(22.14,23)(24.5,14)(19.89,13)(23.41,24)(22.3,46)(31.03,39)(26.83,31)(22.64,39)(19.15,17)(18.13,16)(21.89,20)(24.79,12)(23.59,23)(17.62,40)(24.89,50)(24.04,19)(29.95,17)(29.24,22)(23.93,10)(23.2,9)(20.94,15)(17.69,33)(27.17,43)(25.51,7)(18.74,45)(21.07,26)(23.45,23)(18.46,31)(21.13,16)(21.83,36)(22.45,31)(23.58,23)(30.44,22)(34.38,16)(20.07,30)(23.37,39)(20.26,26)(24.53,14)(25.63,77)(18.5,8)(24.34,8)(22.89,22)(22.49,7)(20.61,13)(19.53,25)(20.02,14)(21.83,34)(24.24,34)(18.28,8)(19.51,23)(25.63,15)(24.71,20)(16.59,52)(21.4,45)(21.36,20)(22.01,18)(20.86,31)(17.49,18)(22.07,31)(15.97,38)(25.81,10)(29.35,5)(20.88,25)(21.51,22)(22.24,10)(19.74,30)(17.63,16)(25.72,33)(21.32,30)(22.06,13)(21.52,25)(19.78,20)(22.19,20)(20.75,2)(24.05,29)(24.23,8)(24.38,26)(20.79,6)(23.72,19)(17.95,19)
      };
      \addlegendentry{Researchers}

    \addplot[
      mark=square,
      ]
      coordinates {
        (15.0,24.7)(17.38,24.27)(19.75,23.83)(22.12,23.4)(24.5,22.96)(26.88,22.53)(29.25,22.09)(31.63,21.66)(34.0,21.22)
      };
      \addlegendentry{Least Squares Fit}

    \end{axis}
    \end{tikzpicture}
    \caption{Positive citation percentage vs. H-index}
    \label{graph:positive-karma}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \begin{tikzpicture}
      \begin{axis}[
          width=(1\textwidth),
          xlabel={Neutral Citation percentage [\%]},
          ylabel={H-index},
          xmin=49, xmax=82,
          ymin=0, ymax=80,
          xtick={50, 55, 60, 65, 70, 75, 80},
          ytick={0,10,20,30,40,50,60,70,80},
          ymajorgrids=true,
          grid style=dashed,
      ]

      \addplot[
        only marks, color=blue, mark size=1.3pt
        ]
        coordinates {
          (70.44,16)(72.45,16)(67.78,5)(68.11,23)(71.3,23)(72.84,22)(59.06,7)(63.2,18)(73.47,25)(66.18,11)(69.69,32)(73.44,26)(69.81,18)(68.2,66)(68.0,40)(69.77,5)(66.48,23)(62.79,14)(64.06,13)(70.05,24)(70.53,46)(63.44,39)(64.29,31)(67.3,39)(69.02,17)(76.45,16)(69.85,20)(69.62,12)(71.86,23)(71.93,40)(59.32,50)(61.88,19)(57.42,17)(63.52,22)(65.97,10)(70.43,9)(64.07,15)(67.43,33)(65.4,43)(67.73,7)(73.79,45)(70.47,26)(70.92,23)(76.83,31)(64.27,16)(63.97,36)(72.25,31)(66.61,23)(65.47,22)(57.82,16)(72.92,30)(68.0,39)(70.04,26)(63.82,14)(67.74,77)(67.91,8)(69.23,8)(73.59,22)(67.3,7)(72.12,13)(72.78,25)(62.61,14)(64.3,34)(63.49,34)(70.75,8)(75.25,23)(68.77,15)(69.08,20)(73.35,52)(72.22,45)(72.52,20)(66.46,18)(73.06,31)(75.99,18)(73.59,31)(80.0,38)(71.16,10)(63.04,5)(75.59,25)(68.07,22)(62.65,10)(71.69,30)(73.76,16)(65.61,33)(65.89,30)(70.13,13)(73.49,25)(74.18,20)(66.88,20)(67.92,2)(67.19,29)(52.16,8)(66.08,26)(59.57,6)(70.01,19)(76.34,19)
        };
        \addlegendentry{Researchers}

      \addplot[
        mark=square,
        ]
        coordinates {
          (50.0,14.63)(53.75,16.4)(57.5,18.17)(61.25,19.94)(65.0,21.71)(68.75,23.48)(72.5,25.24)(76.25,27.01)(80.0,28.78)
        };
        \addlegendentry{Least Squares Fit}

      \end{axis}
      \end{tikzpicture}
      \caption{Neutral citation percentage vs. H-index}
      \label{graph:neutral-karma}
    \end{subfigure}
    \caption{Citation purpose percentages vs. H-index of researchers}
    \label{fig:researcher-karma}
\end{figure*}

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        /pgf/number format/1000 sep={},
        ylabel={Citation Sentiment Percentage Change [\%]},
        xlabel={Year},
        ymin=-10, ymax=30,
        xmin=1999, xmax=2025,
        ytick={-5, 0, 5, 10, 15, 20, 25}, % 40, 50, 60, 70, 80},
        xtick={2000, 2005, 2010, 2015, 2020, 2025},
        ymajorgrids=true,
        grid style=dashed,
        legend style={at={(0.5,-0.5)},anchor=south},
        every axis plot/.append style={very thick}
    ]
    \addplot[
      red
      ]
      coordinates {
        (2000,0.0)(2001,-6.89)(2002,6.41)(2003,14.17)(2004,7.38)(2005,5.46)(2006,-3.63)(2007,-6.46)(2008,4.31)(2009,7.41)(2010,1.63)(2011,10.81)(2012,3.27)(2013,2.26)(2014,11.97)(2015,12.98)(2016,22.32)(2017,12.63)(2018,10.3)(2019,13.03)(2020,19.53)(2021,19.05)(2022,16.14)(2023,26.52)
      };
      \addlegendentry{Negative Citation}

    \addplot[teal]
      coordinates {
        (2000,0.0)(2001,-0.77)(2002,4.82)(2003,1.13)(2004,-1.77)(2005,6.03)(2006,10.46)(2007,8.17)(2008,10.21)(2009,5.34)(2010,4.7)(2011,6.28)(2012,8.05)(2013,8.59)(2014,6.44)(2015,4.86)(2016,9.03)(2017,10.24)(2018,10.57)(2019,8.28)(2020,10.97)(2021,10.02)(2022,14.42)(2023,12.63)
      };
      \addlegendentry{Positive Citation}

    \addplot[gray]
      coordinates {
        (2000,0.0)(2001,0.98)(2002,-2.11)(2003,-1.89)(2004,-0.3)(2005,-2.36)(2006,-2.65)(2007,-1.67)(2008,-3.45)(2009,-2.37)(2010,-1.55)(2011,-3.02)(2012,-2.71)(2013,-2.75)(2014,-3.2)(2015,-2.85)(2016,-5.1)(2017,-4.38)(2018,-4.22)(2019,-3.85)(2020,-5.36)(2021,-5.02)(2022,-5.99)(2023,-6.61)
      };
      \addlegendentry{Neutral Citation}

    \end{axis}
    \end{tikzpicture}
    \caption{Citation sentiment percentage change throughout years}
    \label{graph:year-karma}
\end{figure}

\begin{figure*}
  \centering
  \begin{subfigure}{0.3\textwidth}
    \begin{tikzpicture}
      \begin{axis}[
          width=(1\textwidth),
          ylabel={Negative Citation percentage [\%]},
          xlabel={Researcher's number of papers},
          ymin=0, ymax=30,
          xmin=0, xmax=700,
          ytick={0, 5, 10, 15, 20, 25, 30},
          xtick={0, 100, 200, 300, 400, 500, 600, 700},
          ymajorgrids=true,
          grid style=dashed,
        ]

        \addplot[
          only marks, color=blue, mark size=1.3pt
          ]
          coordinates {
            (34,15.15)(64,10.55)(7,14.44)(99,7.21)(109,9.57)(152,7.83)(27,18.18)(54,8.52)(83,6.66)(31,3.92)(211,9.46)(219,6.8)(62,4.01)(674,11.46)(129,7.98)(8,7.73)(75,11.38)(36,12.7)(55,16.05)(58,6.53)(252,7.17)(88,5.53)(226,8.88)(178,10.06)(75,11.83)(87,5.42)(38,8.27)(40,5.59)(113,4.55)(188,10.45)(155,15.78)(41,14.08)(50,12.64)(67,7.23)(30,10.09)(21,6.36)(40,14.99)(95,14.88)(179,7.43)(22,6.77)(176,7.47)(50,8.47)(106,5.62)(262,4.71)(87,14.61)(215,14.2)(169,5.3)(177,9.81)(79,4.1)(47,7.8)(117,7.02)(128,8.63)(98,9.7)(44,11.65)(493,6.63)(35,13.58)(22,6.42)(65,3.52)(19,10.21)(41,7.27)(170,7.69)(27,17.37)(107,13.87)(126,12.28)(24,10.97)(73,5.24)(29,5.61)(77,6.22)(380,10.06)(229,6.38)(58,6.13)(63,11.54)(107,6.08)(56,6.52)(152,4.33)(393,4.03)(16,3.03)(14,7.61)(70,3.53)(49,10.42)(18,15.11)(228,8.56)(46,8.61)(80,8.67)(124,12.79)(53,7.81)(205,4.99)(104,6.04)(65,10.93)(6,11.32)(109,8.76)(17,23.61)(98,9.54)(17,19.64)(42,6.27)(42,5.71)
          };
          \addlegendentry{Researchers}

        \addplot[
          mark=square,
          ]
          coordinates {
            (0.0,9.74)(86.25,9.25)(172.5,8.76)(258.75,8.27)(345.0,7.78)(431.25,7.29)(517.5,6.8)(603.75,6.31)(690.0,5.82)
          };
          \addlegendentry{Least Squares Fit}

      \end{axis}
    \end{tikzpicture}
    \caption{Negative citation percentage vs. H-index}
    \label{graph:negative-career}
\end{subfigure}
\hfill
\begin{subfigure}{0.3\textwidth}
  \begin{tikzpicture}
    \begin{axis}[
        width=(1\textwidth),
        ylabel={Positive Citation percentage [\%]},
        xlabel={Researcher's number of papers},
        ymin=13, ymax=36,
        xmin=0, xmax=700,
        ytick={15, 20, 25, 30, 35},
        xtick={0, 100, 200, 300, 400, 500, 600, 700},
        ymajorgrids=true,
        grid style=dashed,
    ]

    \addplot[
      only marks, color=blue, mark size=1.3pt
      ]
      coordinates {
        (34,14.41)(64,17.0)(7,17.78)(99,24.68)(109,19.13)(152,19.33)(27,22.76)(54,28.28)(83,19.87)(31,29.9)(211,20.85)(219,19.76)(62,26.18)(674,20.34)(129,24.02)(8,22.5)(75,22.14)(36,24.5)(55,19.89)(58,23.41)(252,22.3)(88,31.03)(226,26.83)(178,22.64)(75,19.15)(87,18.13)(38,21.89)(40,24.79)(113,23.59)(188,17.62)(155,24.89)(41,24.04)(50,29.95)(67,29.24)(30,23.93)(21,23.2)(40,20.94)(95,17.69)(179,27.17)(22,25.51)(176,18.74)(50,21.07)(106,23.45)(262,18.46)(87,21.13)(215,21.83)(169,22.45)(177,23.58)(79,30.44)(47,34.38)(117,20.07)(128,23.37)(98,20.26)(44,24.53)(493,25.63)(35,18.5)(22,24.34)(65,22.89)(19,22.49)(41,20.61)(170,19.53)(27,20.02)(107,21.83)(126,24.24)(24,18.28)(73,19.51)(29,25.63)(77,24.71)(380,16.59)(229,21.4)(58,21.36)(63,22.01)(107,20.86)(56,17.49)(152,22.07)(393,15.97)(16,25.81)(14,29.35)(70,20.88)(49,21.51)(18,22.24)(228,19.74)(46,17.63)(80,25.72)(124,21.32)(53,22.06)(205,21.52)(104,19.78)(65,22.19)(6,20.75)(109,24.05)(17,24.23)(98,24.38)(17,20.79)(42,23.72)(42,17.95)
      };
      \addlegendentry{Researchers}

    \addplot[
      mark=square,
      ]
      coordinates {
        (0.0,22.99)(86.25,22.47)(172.5,21.95)(258.75,21.43)(345.0,20.92)(431.25,20.4)(517.5,19.88)(603.75,19.36)(690.0,18.84)
      };
      \addlegendentry{Least Squares Fit}

    \end{axis}
    \end{tikzpicture}
    \caption{Positive citation percentage vs. H-index}
    \label{graph:positive-career}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \begin{tikzpicture}
      \begin{axis}[
          width=(1\textwidth),
          ylabel={Neutral Citation percentage [\%]},
          xlabel={Researcher's number of papers},
          ymin=49, ymax=82,
          xmin=0, xmax=700,
          ytick={50, 55, 60, 65, 70, 75, 80},
          xtick={0, 100, 200, 300, 400, 500, 600, 700},
          ymajorgrids=true,
          grid style=dashed,
      ]

      \addplot[
        only marks, color=blue, mark size=1.3pt
        ]
        coordinates {
          (34,70.44)(64,72.45)(7,67.78)(99,68.11)(109,71.3)(152,72.84)(27,59.06)(54,63.2)(83,73.47)(31,66.18)(211,69.69)(219,73.44)(62,69.81)(674,68.2)(129,68.0)(8,69.77)(75,66.48)(36,62.79)(55,64.06)(58,70.05)(252,70.53)(88,63.44)(226,64.29)(178,67.3)(75,69.02)(87,76.45)(38,69.85)(40,69.62)(113,71.86)(188,71.93)(155,59.32)(41,61.88)(50,57.42)(67,63.52)(30,65.97)(21,70.43)(40,64.07)(95,67.43)(179,65.4)(22,67.73)(176,73.79)(50,70.47)(106,70.92)(262,76.83)(87,64.27)(215,63.97)(169,72.25)(177,66.61)(79,65.47)(47,57.82)(117,72.92)(128,68.0)(98,70.04)(44,63.82)(493,67.74)(35,67.91)(22,69.23)(65,73.59)(19,67.3)(41,72.12)(170,72.78)(27,62.61)(107,64.3)(126,63.49)(24,70.75)(73,75.25)(29,68.77)(77,69.08)(380,73.35)(229,72.22)(58,72.52)(63,66.46)(107,73.06)(56,75.99)(152,73.59)(393,80.0)(16,71.16)(14,63.04)(70,75.59)(49,68.07)(18,62.65)(228,71.69)(46,73.76)(80,65.61)(124,65.89)(53,70.13)(205,73.49)(104,74.18)(65,66.88)(6,67.92)(109,67.19)(17,52.16)(98,66.08)(17,59.57)(42,70.01)(42,76.34)
        };
        \addlegendentry{Researchers}

      \addplot[
        mark=square,
        ]
        coordinates {
          (0.0,67.26)(86.25,68.27)(172.5,69.28)(258.75,70.29)(345.0,71.3)(431.25,72.31)(517.5,73.32)(603.75,74.33)(690.0,75.34)
        };
        \addlegendentry{Least Squares Fit}

      \end{axis}
      \end{tikzpicture}
      \caption{Neutral citation percentage vs. H-index}
      \label{graph:neutral-career}
    \end{subfigure}
    \caption{Citation purpose percentages vs. researchers' number of papers}
    \label{fig:researcher-career}
\end{figure*}

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}[
        ybar=10pt,
        bar width=15pt,
        enlarge x limits=0.25,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={Percentage of Citation Purposes [\%]},
        symbolic x coords={Positive,Neutral,Negative},
        xtick=data,
        nodes near coords={\pgfmathprintnumber\pgfplotspointmeta\%},
        nodes near coords align={vertical},
        ]
    \addplot coordinates {(Positive,25.1) (Neutral,67.4) (Negative,7.5)};
    \addplot coordinates {(Positive,22.5) (Neutral,68.6) (Negative,8.9)};
    \legend{Same Institution,Different Institution}
    \end{axis}
    \end{tikzpicture}
    \caption{Citation purpose percentages within and outside of the institution}
    \label{graph:institute-karma}
\end{figure}

We see a distribution of 21.59\% positive, 65.58\% neutral, 8.47\% negative, and 4.37\% bad context in Figure \ref{graph:distribution}. This data might not represent the actual distribution of citation sentiments in computer science because all the papers being cited belongs to University of Waterloo professors. Given that University of Waterloo has a relatively high ranking in computer science it is possible that the distribution of citation sentiments is skewed towards positive and neutral.
